Interactively Visualizing Multivariate Market Segmentation Using the R Package Lionfish

### 2024-10-01 manuscript.tex
tidied up heatmap sections more

The 2D tour view can be used to understand the sensitivity of the grouping to individual features. When using the manual tour to remove one feature at a time, the relationship between this variable and the partition between clusters can be examined. The ideal starting view is that obtained through a projection pursuit guided tour~\citep{CBCH94} that optimized for the separation between the labeled groups~\citep{lckl2005}.

@@

where $c_{ij}, i=1, ..., p$ (number of features); $j=1, ..., k$ (number of clusters) are a summary of each feature in each cluster. Because our examples use binary features,  the $c_{ij}$ is the number of 1's of feature $j$ in cluster $i$. %For this setting we have developed a heatmap display that we have found to be particularly useful in feature selection. The idea is to understand differences in the binary ratings between clusters, where we use different normalizations to better explore the data. For \(c_{ij}\) the count (number of true values in the binary feature) in the \(i\)-th cluster and \(j\)-th feature, we define
There are several ways that these values can be normalized to examine different aspects: $f_{ij}^{o} = \frac{c_{ij}}{n}$, $f_{ij}^{c} = \frac{c_{ij}}{n_{i}}$ and $f_{ij}^{f} = \frac{c_{ij}}{n_{j}}$. The first is the overall fraction, where counts are normalised by the overall number of observations $n$. It is generally not particularly useful. The second, $f^c$ is normalized relative to the size of each cluster, and is useful for comparing the features which contribute most to each cluster. The last, $f_f$ is normalized relative to each feature, so it is useful to examine how features are distributed across clusters \textcolor{red}{Not sure this last explanation is good - at least I don't understand what I've written}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winteractiv_heatmap.png}
    \caption{Traditional overview of clusters. Color represents (A) the intra-cluster fraction, and (B) the intra-feature fraction, with lighter indicating higher values. From A, we can see that some activities dominate most clusters, e.g. relaxing and going for walks. Some activities are unique to clusters, e.g. skiing is a primary activity for clusters 1, 3, 5, 6. From B, we can see which activities dominate each cluster, e.g. cluster 4 tourists like going to a spa and using health activities.}
    \label{fig:winteractiv_heatmap}
\end{figure}

\textcolor{red}{It might be better to standardise these intra fractions. For A, one should divide by cluster count, and then compute activity fraction. For B, one should divide by activity count and then compute cluster fraction.} In Figure \ref{fig:winteractiv_heatmap}A, we can observe the general interests of tourists within each cluster. For instance, tourists in clusters 1, 3, 5, and 6 predominantly engaged in alpine skiing, while those in clusters 2 and 4 did not. Additionally, we see that activities such as ski touring and horseback riding were generally unpopular across all clusters.

In Figure \ref{fig:winteractiv_heatmap}B, we can determine whether tourists in a particular cluster had a strong preference for certain activities. For example, nearly all tourists who visited museums are grouped in cluster 2, and those who used health facilities are primarily attributed to cluster 4. We can also identify activities that were similarly popular across all clusters, such as relaxing.

By using the heatmap with these metrics, we can perform feature selection by removing unpopular activities and those that were consistently similar across all clusters. After performing the feature selection by unchecking the corresponding checkboxes in the GUI using this strategy, the following 12 activities remained: alpine skiing, going to a spa, using health facilities, hiking, going for walks, excursions, going out in the evening, going to discos/bars, shopping, sightseeing, museums, and pool/sauna.

>>>>

The 2D tour view can be used to understand the sensitivity of the grouping to individual features. When using the manual tour to remove one feature at a time, the relationship between this variable and the partition between clusters can be examined. The ideal starting view is that obtained through a projection pursuit guided tour~\citep{CBCH94} that was optimized for the separation between the labeled groups~\citep{lckl2005}.

@@

where $c_{ij}, i=1, ..., k$ (number of clusters); $j=1, ..., p$ (number of features) are a summary of each feature in each cluster. Because our examples use binary features,  the $c_{ij}$ is the number of 1's of feature $j$ in cluster $i$. %For this setting we have developed a heatmap display that we have found to be particularly useful in feature selection. The idea is to understand differences in the binary ratings between clusters, where we use different normalizations to better explore the data. For \(c_{ij}\) the count (number of true values in the binary feature) in the \(i\)-th cluster and \(j\)-th feature, we define
There are several ways that these values can be normalized to examine different aspects: $f_{ij}^{o} = \frac{c_{ij}}{n}$, $f_{ij}^{c} = \frac{c_{ij}}{n_{i}}$ and $f_{ij}^{f} = \frac{c_{ij}}{n_{j}}$, where $n_i, n_j$ are the row and column totals. The first, $f^{o}$ is the overall fraction, where counts are normalised by the overall number of observations $n$. Is generally less useful.

The second, $f^c$ is normalized relative to the size of each cluster and can be considered the distribution of features in each cluster. It is useful for comparing the composition of each cluster relative to the features. For example, for $p=4$, if $f^c_{1j} = (0.4, 0.4, 0.1, 0.1)$ it suggests that high values (1's) of features 1 and 2 distinguish cluster 1. In the context of the later examples, this would mean cluster 1 contains tourists that list engaging in activities 1 and 2 but not 3 and 4. This normalisation produces what is called the {\bf intra-cluster fraction}.

The last, $f_f$ is normalized relative to each feature, which can be considered to be the distribution of clusters on each feature. It is useful to examine how features are related to a cluster. Considering the last metric, for example, if $f^f_{i1} = (0.5, 0.4, 0, 0.1, 0)$ ($k=5$) would indicate that high values on feature 1 primarily are in clusters 1, 2. In the context of the later examples, this would mean activity 1 is commonly listed only in clusters 1 and 2. This normalisation produces what is called the {\bf intra-feature fraction}.

Generally, if the features are ordered categories or numerical scores, decomposing the cluster-feature summary is still possible by using an analysis of variance breakdown into row and column effects.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winteractiv_heatmap.png}
    \caption{Traditional overview of clusters. Color represents (A) the intra-cluster fraction, and (B) the intra-feature fraction, with lighter indicating higher values. From A, we can see that cluster 3 tourists like alpine skiing, going out in the evening and going to discos and bars. They also like relaxing, shopping and sightseeing but these are popular among all tourists. From B, we can see the distribution of activities across clusters, e.g. using health facilities is popular in cluster 4 while going to a pool or sauna is popular in clusters 1 and 5.}
    \label{fig:winteractiv_heatmap}
\end{figure}

In Figure \ref{fig:winteractiv_heatmap}A, where colors show normalised column counts, we can observe the general interests of tourists within each cluster. Some activities are high on almost all clusters (e.g. relaxing, shopping), and some are low on all clusters (e.g. ski touring and horseback riding), and ignoring these can be helpful when assessing the distribution of activities between clusters. When comparing clusters 5 and 6 we can see that alpine skiing is high in both, but cluster 5 tourists also like going to the pool or sauna, but cluster 6 tourists prefer going for walks.  %For instance, tourists in clusters 1, 3, 5, and 6 predominantly engaged in alpine skiing, while those in clusters 2 and 4 did not. Additionally, we see that activities such as ski touring and horseback riding were generally unpopular across all clusters.

In Figure \ref{fig:winteractiv_heatmap}B, we can determine whether for a particular feature by tourists in which clusters.  For example, nearly all tourists who visited museums are grouped in cluster 2, and those who used health facilities are primarily attributed to cluster 4. Some activities, such as relaxing, are popular across all clusters.

These heatmaps can help with selecting features to focus on using the tour. Unpopular and universally popular activities can be removed. After performing the feature selection by unchecking the corresponding checkboxes in the GUI using this strategy, the following 12 activities remained: alpine skiing, going to a spa, using health facilities, hiking, going for walks, excursions, going out in the evening, going to discos/bars, shopping, sightseeing, museums, and pool/sauna.
